### 2021-01-12 paper/motivating_results.tex

\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{graphicx,tikz}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage{bm,fixmath}
\usepackage{color}
\usepackage{subcaption}
\usepackage{xspace}

\usetikzlibrary{calc,shapes,arrows.meta,decorations.pathreplacing,decorations.markings}

% math operators
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\conv}{conv}
\DeclareMathOperator{\prox}{prox}
\DeclareMathOperator{\proj}{proj}
\DeclareMathOperator{\fix}{fix}
\DeclareMathOperator{\diam}{diam}
\DeclareMathOperator{\image}{Im}
\DeclareMathOperator{\dist}{dist}
\DeclareMathOperator{\udist}{\overline{dist}}
\DeclareMathOperator{\ldist}{\underline{dist}}

\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
% double inner product
\newcommand{\zprod}[1]{\langle\kern-0.25ex\langle#1\rangle\kern-0.25ex\rangle}
% triple norm
\newcommand{\znorm}[1]{{\left\vert\kern-0.25ex\left\vert\kern-0.25ex\left\vert #1 
    \right\vert\kern-0.25ex\right\vert\kern-0.25ex\right\vert}}

% theorem environments
\usepackage{amsthm}

\theoremstyle{plain} % Plain: italic text, extra space above and below
\newtheorem{example}{Example}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{axiom}[theorem]{Axiom}

\theoremstyle{definition} % Definition: upright text, extra space above and below
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{conjecture}[theorem]{Conjecture}

\theoremstyle{remark} % Remark: upright text, no extra space above or below
\newtheorem{remark}[theorem]{Remark}

% algorithms packages and commands
\usepackage{algorithm}
\usepackage{algpseudocode}

% define commands for algorithm environment
\algrenewcommand\algorithmicrequire{\textbf{Input:}}

% new math commands
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}

\newcommand{\x}{\mathbold{x}}
\newcommand{\y}{\mathbold{y}}

\newcommand{\I}{\mathcal{I}}
\newcommand{\T}{\mathcal{T}}

% file's info
\author{Nicola Bastianello}
\title{Motivating Results}

\begin{document}

\maketitle

%---------------------------------------------------------------------------
\section{Result 1 (static)}
The following result shows that the fixed point residual converges linearly fast for contractive operators, but sub-linearly (with the square root of the iteration number) for averaged operators.

\begin{proposition}
Consider an operator $\T : \R^n \to \R^n$ and the Banach-Picard iteration
$$
	\x^{\ell+1} = \T \x^\ell, \quad \ell \in \N.
$$
Then, the fixed point residual $\{ \norm{\x^{\ell+1} - \x^\ell} \}$ is bounded as:
\begin{itemize}
	\item \emph{$\T$ $\zeta$-contractive}: $\norm{\x^{\ell+1} - \x^\ell} \leq (1 + \zeta) \zeta^\ell \norm{\x^0 - \x^*}$;

	\item \emph{$\T$ $\alpha$-averaged}: $\norm{\x^{\ell+1} - \x^\ell} \leq \frac{1}{\sqrt{\ell+1}} \sqrt{\frac{\alpha}{1 - \alpha}} \norm{\x^0 - \x^*}$.
\end{itemize}
\end{proposition}
%
\begin{proof}
\emph{$\T$ contractive}: by \cite[Theorem~1.50]{bauschke_convex_2017} we know that
$$
	\norm{\x^\ell - \x^*} \leq \zeta^\ell \norm{\x^0 - \x^*};
$$
using the triangle inequality we then have
\begin{align*}
	\norm{\x^{\ell+1} - \x^\ell} &\leq \norm{\x^{\ell+1} - \x^*} + \norm{\x^\ell - \x^*} \\
	&\leq \zeta \norm{\x^\ell - \x^*} + \norm{\x^\ell - \x^*} \\
	&\leq (1 + \zeta) \norm{\x^\ell - \x^*} \\
	&\leq (1 + \zeta) \zeta^\ell \norm{\x^0 - \x^*}.
\end{align*}

\emph{$\T$ averaged}: by the averagedness of $\T$ we can write, for some $\x^* \in \fix(\T)$:
$$
	\norm{\x^{\ell+1} - \x^*}^2 \leq \norm{\x^\ell - \x^*}^2 - \frac{1 - \alpha}{\alpha} \norm{\x^{\ell+1} - \x^\ell}^2;
$$
rearranging, and summing over time we get
$$
	\sum_{h = 0}^\ell \norm{\x^{h+1} - \x^h}^2 \leq \frac{\alpha}{1 - \alpha} \norm{\x^0 - \x^*}^2
$$
where we used the telescoping sum and removed the negative term $- \norm{\x^{\ell+1} - \x^\ell}^2$.

Since $\{ \norm{\x^{h+1} - \x^h}^2 \}$ is a monotonically decreasing sequence, we know that $(\ell+1) \norm{\x^{\ell+1} - \x^h} \leq \sum_{h = 0}^\ell \norm{\x^{h+1} - \x^h}^2$; using this fact, rerranging and taking the square root yields the thesis.
\end{proof}

%---------------------------------------------------------------------------
\section{Result 2 (dynamic)}
Consider the time-varying Banach-Picard
$$
	\x_{k+1} = \T_{k+1} \x_k, \quad k \in \N,
$$
and for simplicity assume that $\x_0 \in \fix(\T_0)$.

\begin{proposition}
The cumulative fixed point residual is bounded as:
\begin{itemize}
	\item \emph{$\T_k$ $\zeta$-contractive}:
	$$
		\frac{1}{k+1} \sum_{h = 0}^k \norm{\x_{h+1} - \x_h}^2 \leq \left( \sigma \frac{1 + \zeta}{1 - \zeta} \right)
	$$
	where $\sigma \geq \max_k \norm{\x_k^* - \x_{k-1}^*}$;

	\item \emph{$\T_k$ $\alpha$-averaged} and defined on a compact domain:
	$$
		\frac{1}{k+1} \sum_{h = 0}^k \norm{\x_{h+1} - \x_h}^2 \leq \frac{\alpha}{1 - \alpha} (\tau^2 + 2 \delta \tau)
	$$
	where $\tau$ is the maximum distance between consecutive fixed point sets $\fix(\T_k)$ and $\fix(\T_{k-1})$, and $\delta$ is the diameter of $\T_k$'s domains.
\end{itemize}
\end{proposition}
%
\begin{proof}
\emph{$\T_k$ contractive}: using triangle inequality and contractiveness we can write
\begin{align*}
	\norm{\x_{k+1} - \x_k} &\leq \norm{\x_{k+1} - \x_{k+1}^*} + \norm{\x_k - \x_{k+1}^*} \\
	&\leq (1 + \zeta) \norm{\x_k - \x_{k+1}^*} \\
	&\leq (1 + \zeta) (\norm{\x_k - \x_k^*} + \sigma) \\
	&\leq (1 + \zeta) \left( \zeta^k \norm{\x_0 - \x_0^*} + \frac{\sigma}{1 - \zeta} \right) = \frac{1 + \zeta}{1 - \zeta} \sigma.
\end{align*}

Taking the square and averaging over time we get:
\begin{align*}
	\frac{1}{k+1} \sum_{h = 0}^k \norm{\x_{h+1} - \x_h}^2 \leq \left( \frac{1 + \zeta}{1 - \zeta} \sigma \right)^2.
\end{align*}

\emph{$\T_k$ averaged}: using the averagedness we have
$$
	\norm{\x_{k+1} - \x_{k+1}^*}^2 \leq \norm{\x_k - \x_{k+1}^*}^2 - \frac{1 - \alpha}{\alpha} \norm{\x_{k+1} - \x_k}^2;
$$
rearranging and using triangle and Cauchy-Schwarz inequalities we have
\begin{align*}
	\norm{\x_{k+1} - \x_k}^2 &\leq \frac{\alpha}{1 - \alpha} \left( \norm{\x_k -\x_{k+1}^*}^2 - \norm{\x_{k+1} - \x_{k+1}^*}^2 \right) \\
	&\leq \frac{\alpha}{1 - \alpha} \left( \norm{\x_k - \x_k^*}^2 - \norm{\x_{k+1} - \x_{k+1}^*}^2 + \norm{\x_k^* - \x_{k+1}^*}^2 + 2 \norm{\x_k - \x_{k+1}^*} \norm{\x_k^* - \x_{k+1}^*} \right) \\
	&\leq \frac{\alpha}{1 - \alpha} \left( \norm{\x_k - \x_k^*}^2 - \norm{\x_{k+1} - \x_{k+1}^*}^2 + \tau^2 + 2 \delta \tau \right)
\end{align*}
where the last inequality follows by $\norm{\x_k^* - \x_{k-1}^*} \leq \tau$ and $\delta$ being the diameter of the compact domain.

Averaging over time, and noticing that the error terms cancel out we have
$$
	\frac{1}{k+1} \sum_{h = 0}^k \norm{\x_{h+1} - \x_h}^2 \leq \frac{\alpha}{1 - \alpha} \left( \frac{1}{k+1} \norm{\x_0 - \x_0^*}^2 + \tau^2 + 2 \delta \tau \right) = \frac{\alpha}{1 - \alpha} (\tau^2 + 2 \delta \tau).
$$
\end{proof}

\bibliographystyle{ieeetr}
\bibliography{references}

\end{document}